Deep Gaussian Processes and Hierarchical Modeling with Functional Data: A Case Study in Cosmological Power Spectra

### 2024-05-24 paper/main.tex
add initial pred figures

\oddsidemargin=0.25in
\evensidemargin=0.25in
\textwidth=7in
\textheight=8.75in
\topmargin=-.5in
\addtolength{\oddsidemargin}{-.5in}
\addtolength{\evensidemargin}{-.5in}
\footskip=0.5in
%\doublespacing

@@

\textcolor{blue}{The Cosmic Emu emulator is constructed by (process convolutions, ...). Segue to our method after and compare/contrast with Cosmic Emu.}

Due to the nonstationary behavior expected to be seen in power spectra (e.g., from baryonic acoustic oscillations), we expand on the previous emulation methods and propose a novel synthesis of methods to quantify uncertainty and predict matter power spectra for different cosmologies using deep Gaussian processes (DGPs). In order to handle multiple realizations of functional output (from the Mira-Titan simulation suite), we use hierarchical and basis functions modeling within our DGP framework to accurately estimate and predict power spectra for different cosmologies. When comparing to Cosmic Emu, we see favorable results for our method. The code to reproduce these results are available at \url{https://github.com/stevewalsh124/dgp.hm}. \textcolor{blue}{Please note: this repository is currently private until we hear back regarding permissions for sharing the Mira-Titan data alongside our code.}

@@

Here we show results on held-out cosmologies.  We compare to cosmicEMU and Jared's convolution approach. We can also discuss estimation of main effects, interactions amongst the 8 parameters, and variance decomposition.

\begin{itemize}
    \item compare the 4 with boxplot: our DGP training, DGP testing, vs proc conv training (aka Jared's files), proc conv testing (aka Cosmic EMU)
    \item compare the 4's RMSEs as a function of $k$
    \item Main effects
\end{itemize}

>>>>

\oddsidemargin=0.25in
\evensidemargin=0.25in
\textwidth=7in
\textheight=8.75in
\topmargin=-.5in
\addtolength{\oddsidemargin}{-.5in}
\addtolength{\evensidemargin}{-.5in}
\footskip=0.5in
\doublespacing

@@

In this work, we expand on the previous emulation methods and propose a novel synthesis of methods to quantify uncertainty and predict matter power spectra for different cosmologies using deep Gaussian processes (DGPs). In order to handle multiple realizations of functional output (from the Mira-Titan simulation suite), we use hierarchical and basis functions modeling within our DGP framework to accurately estimate and predict power spectra for different cosmologies. When comparing to Cosmic Emu, we see favorable results for our method. The code to reproduce these results are available at \url{https://github.com/stevewalsh124/dgp.hm}. \textcolor{blue}{Please note: this repository is currently private until we hear back regarding permissions for sharing the Mira-Titan data alongside our code.}

@@

Here we show results on held-out cosmologies.  The Cosmic Emu emulator is constructed by (process convolutions, ...). Segue to our method after and compare/contrast with Cosmic Emu.

In addition to predicting for held-out cosmologies, we can also predict over a large uniform or full factorial design to obtain estimated main effects for each parameter in $\theta$, interactions amongst the 8 parameters, and variance decomposition. Should this be included?

\begin{figure}[ht]
    \centering
    \includegraphics[width=6in]{rmse_by_k.png}
    \caption{Average RMSE of each method across all $k$ values.}
    \label{fig:plot_rmse_k}
\end{figure}

\begin{figure}[ht]
    \centering
    \includegraphics[width=6in]{rmse_box.png}
    \caption{Boxplot of all RMSEs of each method across all six test models. From top to bottom: box plots for our training method, process convolution training method, Cosmic EMU predictions, and our DGP-PC predictions.}
    \label{fig:plot_rmse_box}
\end{figure}
